---
title: "Assignment2_XinHuang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, results='hide', include=FALSE}
library("plyr")
library("dplyr")
library("tidyr")
library("lattice")
library(rgl)
library(rglwidget)
```

#Answer for Question 1  
Run the following code to generate the figures:

```{r echo = TRUE, fig=TRUE, fig.align='center'}
#set workspace and read file
setwd("/Users/xinhuang/Google Drive/CSC 522 R Language Programming /Homework2")
sfData <- read.csv("AirQualitySanFranciscoData.csv", header = TRUE)
sdData <- read.csv("AirQualitySanDiegoData.csv", header = TRUE)

# set layout
mat <- matrix(1 : 4, nrow = 2)
layout(mat)

with(sfData, {
    plot(Year, ExceedStateLevelDays1.hr, col = "red", ylim = c(0, 100), 
         type = "b", ylab = "Number of Days over Standard", 
         pch = 20, lty = 2)
    points(Year, ExceedStateLevelDays8.hr, col = "yellow", 
           type = "b", pch = 20, lty = 3)
    points(Year, ExceedNationalLevelDays8.hr, col = "green", 
           type = "b", pch = 20, lty = 4)
    title(main = "San Francisco Air Quality")
    legend("topright", 
           c("State Days 1 hour", "State Days 8 hours", "National Level 8 hours"),
           col = c("red", "yellow", "green"), 
           pch = 1, cex = 0.5, title = "Over standard")
})

with(sfData, {
    plot(Year, MaxConcentration1.hr, col = "cyan", ylim = c(0.04, 0.17), 
         type = "b", ylab = "Maximum Concentration", pch = 20, lty = 2)
    points(Year, MaxConcentration8.hr, col = "blue", 
           type = "b", pch = 20, lty = 3)
    legend("topright", c("Max 1 hour", "Max 8 hour"),
           col = c("cyan", "blue"), pch = 1, cex = 0.5, title = "ppm")
})

with(sdData, {
    plot(Year, ExceedStateLevelDays1.hr, col = "red", ylim = c(0, 100), 
         type = "b", ylab = "Number of Days over Standard", 
         pch = 20, lty = 2)
    points(Year, ExceedStateLevelDays8.hr, col = "yellow", 
           type = "b", pch = 20, lty = 3)
    points(Year, ExceedNationalLevelDays8.hr, col = "green", 
           type = "b", pch = 20, lty = 4)
    title(main = "San Diego Air Quality")
    legend("topright", 
           c("State Days 1 hour", "State Days 8 hours", "National Level 8 hours"),
           col = c("red", "yellow", "green"), 
           pch = 1, cex = 0.5, title = "Over standard")
})

with(sdData, {
    plot(Year, MaxConcentration1.hr, col = "cyan", ylim = c(0.04, 0.17), 
         type = "b", ylab = "Maximum Concentration", 
         pch = 20, lty = 2)
    points(Year, MaxConcentration8.hr, col = "blue", 
           type = "b", pch = 20, lty = 3)
    legend("topright", c("Max 1 hour", "Max 8 hour"),
           col = c("cyan", "blue"), pch = 1, cex = 0.5, title = "ppm")
})

mtext("Main title", line=2, font=2, cex=1.2, outer = TRUE)
```

#Answer for Question 2  
Run the following code to generate the figures:  

```{r echo = TRUE, fig=TRUE, fig.align='center'}
#Question 2
myState <- as.data.frame(state.x77)
d <- density(myState$Area)
axisV <- as.integer(c(min(myState$Area),
                      mean(myState$Area),
                      myState[rownames(myState)== "Texas",
                              colnames(myState)=="Area"],
                      max(myState$Area))) 

plot(d, col = "red", lwd = 2,
    xlim=c(min(myState$Area),max(myState$Area)),
    xlab = "Land Area in Square Miles",
    ylab = "Density",
    xaxt = "n",
    main = "Probability Density Plot \n State Land Area in Square Miles")
axis(1, at = axisV, labels = axisV) 
abline(v=mean(myState$Area),col = "black", lty = 5) 
text(2 * mean(myState$Area) + 20000, 11e-06,
     paste("mean = ",round(mean(myState$Area), digits=0)))
text(myState$Area[rownames(myState)=="Alaska"]-85000, 8e-06,
     paste("Alaska Area = \n",myState$Area[rownames(myState)=="Alaska"])) 
text(myState$Area[rownames(myState)=="Texas"]-3000, 8e-06,
     paste("Texas Area = \n",myState$Area[rownames(myState)=="Texas"]))

hist(myState$Area, 
     breaks= 50, 
     col="red",
     xaxt = "n",
     xlim=c(min(myState$Area),max(myState$Area)),
     xlab = "Land Area in Square Miles",
     main = "Histogram of State Land Area in Square Miles")
axis(1, at = axisV, labels = axisV) 
```

#Answer for Question 3  

```{r fig=TRUE, fig.align='center'}
#Question 3
##Figure 1
myState <-as.data.frame(cbind(state.x77, region = state.region)) 
myState <- cbind(myState, regionName = levels(state.region)[state.region]) 

stateSum <- myState %>% 
    group_by(regionName) %>% 
    summarise_all(sum)
target <- c("Northeast", "South", "North Central", "West")
stateSum <- stateSum[match(target, stateSum$regionName), ]

barplot(stateSum$Area,
        names.arg = stateSum$regionName,
        col = c("red", "yellow", "green", "blue"),
        xlab = "Region",
        ylab = "Area in Square Miles",
        ylim = c(0, 1700000),
        main = "Area By Region")

##Figure 2
myState <- myState[order(myState$regionName, myState$Area),] 
myState$regionName <- factor(myState$regionName) 
myState$color[myState$regionName == "Northeast"] <- "red"
myState$color[myState$regionName == "South"] <- "yellow"
myState$color[myState$regionName == "North Central"] <- "green"
myState$color[myState$regionName == "West"] <- "blue"

myState$Area[rownames(myState)=="Texas"]

TexasArea <- myState$Area[rownames(myState)=="Texas"]
AlaskaArea <- myState$Area[rownames(myState)=="Alaska"]

dotchart(myState$Area, 
         labels= rownames(myState), 
         cex= 1,
         groups = myState$regionName,
         gcolor = "black",
         color = myState$color,
         xlim = c(min(myState$Area), 170000),
         main = "Area in Square Miles \n grouped by region",
         xlab ="Area of State in Square Miles",
         pch = 20)
mtext(paste("Texas Area = ", TexasArea), side = 1, line= 2, adj = 0, col = "yellow", cex = .5)
mtext(paste("Alaska Area = ", AlaskaArea), side = 1, line= 2, adj = 1, col = "blue", cex = .5)

#Figure 3
d1 <- density(myState[which(myState$regionName == "Northeast"),]$Area)
d2 <- density(myState[which(myState$regionName == "South"),]$Area)
d3 <- density(myState[which(myState$regionName == "North Central"),]$Area)
d4 <- density(myState[which(myState$regionName == "West"),]$Area)

plot(d1, col = "red", lwd = 2,
     xlim = c(-10000, 750000),
     xlab = "Area of States",
     ylim = c(0, 0.000030),
     ylab = "Density",
     lty = 2, main = "")
box(lty = 1, col = 'black')
lines(d2, col = "yellow", lty = 3)
lines(d3, col = "green", lty = 1)
lines(d4, col = "blue", lty = 4)
legend("topright", c("Northeast", "North Central", "South", "West"),
       col = c("red", "green", "yellow", "blue"), 
       lty = 1, cex = 0.8,
       title = "Region Names")
mtext("Probability Density Comparison of Areas by Regions", line = 1, outer = FALSE)
```

#Answer for Question 4  
Run the following code to generate the figures:  

```{r fig=TRUE, fig.align='center'}
##Question 4
myState <-as.data.frame(cbind(state.x77, region = state.region)) 
myState <- cbind(myState, regionName = levels(state.region)[state.region])
myState <- myState[order(myState$regionName, myState$Area),] 
myState$regionName <- factor(myState$regionName) 
myState$color[myState$regionName == "Northeast"] <- "red"
myState$color[myState$regionName == "South"] <- "yellow"
myState$color[myState$regionName == "North Central"] <- "green"
myState$color[myState$regionName == "West"] <- "blue"
##Figure 1
par(mar = c(4, 4, 3, 5), oma = c(0, 0, 0, 0)) 
pairs(myState[,2:6], main="Pairs Plot", pch = 21,
      bg = c("green", "red", "yellow", "blue")[unclass(myState$region)])

##Figure 2
mat <- matrix(1 : 4, nrow = 2)
layout(mat)

plot(myState$Illiteracy,  myState$`Life Exp`, 
     col = myState$color, pch = 20, bg = myState$color,
     xlab = "Illiteracy",
     ylab = "Life Exp",
     main = "Illiteracy vs Life Expectancy")

legend("topright", 
       legend = c("Region Names", "Northeast", "North Central", "South", "West"),
       col = c("blue","red", "green", "yellow", "blue"),
       lty = 1, cex = 0.2,
       title = "Region Names")

plot(myState$Income,  myState$`Life Exp`, 
     col = myState$color, pch = 20, bg = myState$color,
     xlab = "Incoume",
     ylab = "Life Exp",
     main = "Income vs Life Expectancy")

plot(myState$Illiteracy,  myState$`HS Grad`, 
     col = myState$color, pch = 20, bg = myState$color,
     xlab = "Illiteracy",
     ylab = "HS Grad",
     main = "Illiteracy vs Illiteracy vs High School Graduation")

plot(myState$Income,  myState$`HS Grad`, 
     col = myState$color, pch = 20, bg = myState$color,
     xlab = "Income",
     ylab = "HS Grad",
     main = "Illiteracy vs Income vs High School Graduation")
```

You can drive the following observations:  
1. The higher education one has, the higher income he might have  
2. The lower percentage of one region, the longer time spane of that region can have

#Answer for Question 5  
##Using precip dataset

```{r fig=TRUE, fig.align='center'}
## dotchart
require(graphics)
dotchart(precip[order(precip)], main = "precip data",
        pch = 20, cex = 0.5, color = "firebrick4")
title(sub = "Average annual precipitation")
```

##Using state.x77 dataset  
Use $plot3d()$ to plot column $Income , Illiteracy, Life Exp$
```{r fig=TRUE, fig.align='center'}
myState <-as.data.frame(cbind(state.x77, region = state.region)) 
attach(myState)
mfrow3d(nr = 1, nc = 1, sharedMouse = TRUE) 
plot3d(Income , Illiteracy, myState$`Life Exp`, col="red", size=10)
rglwidget()
```


